Predictive modeling using genomic annotations

### 2018-11-26 Manuscript/manuscript.Rmd
manuscript edits

##Variable Reduction

We evaluated the use of two variable reduction algorithms, the LASSO and Elastic-Net regularizations. Each algorithm was performed prior to re-sampling. For the LASSO, the $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10. The additional $\alpha$ term for the Elastic-Net was tuned

@@

Figure 3 details the ensemble framework proposed and evaluated in this study. For each resolution and predictor type chosen, the data, composed of majority and minority classes, was split into a 7:3 ratio of training to testing sets. For the training set, a variable reduction technique was performed. The feature space was reduced and a resampling technique was applied. A classification algorithm in the form of a random forest was performed and validated using the test set. Details of each step in the framework are provided below.

@@

@@

@@

   4.	SMOTE (Synthetic Minority Over-Sampling Technique): This method incorporated both random under- and over-sampling. Under-sampling is performed without replacement from the majority class, while over-sampling is performed by creating new synthetic observations using the minority class. The SMOTE algorithm was devised by Chawla et al and stored in the DMwR package, available in R. The user is able to specify the parameters denoting the percent of over- and under-sampling from the two classes. This allows the user to control the proportion of disbalance in the data. For the purposes of this paper, we used SMOTE to create perfectly balanced data. However further research would be needed to see if varying the proportion of class disbalance could improve predictive performance.

Figure 3 details the ensemble framework proposed and evaluated in this study. For each resolution and predictor type chosen, the data, composed of majority and minority classes, was split into a 7:3 ratio of training to testing sets. For the training set, a variable reduction technique was performed. The feature space was reduced and a resampling technique was applied. A classification algorithm in the form of a random forest was performed and validated using the test set. 10-fold cross validation was used such that each sub-training set was partitioned into a fixed 9:1 ratio.  The model was performed and assessed using the holdout set to get more stable performance metrics and reduce bias.

@@

>>>>

@@

Figure 3 details the ensemble framework proposed and evaluated in this study. For each resolution and predictor type chosen, the data was split into training to testing sets. A variable reduction technique was then performed. The feature space was reduced and a resampling technique was applied. Lastly, a classification algorithm in the form of a random forest was performed and validated. Details of each step in the framework are provided below.

@@

Given a particular resolution, the full data set was composed of genomic bins that either overlapped with a flanked TAD boundary (minority set) or did not (majority set). Figure 3 highlights the severe class imbalance present in both the full data set. The data was then split into a 7:3 ratio of training to testing sets. Each training and testing set was composed of a similar majority to minority class ratio as the full data set, which is represented by the segmented regions in Figure 3. The training set was then used to perform all downstream modeling. The same test set was held separate and used to validate all models.

@@

We evaluated the use of two variable reduction algorithms, the LASSO and Elastic-Net regularizations. Each algorithm was performed prior to re-sampling. For the LASSO, the $\lambda$ penalization term was tuned over a grid of 10 values on an exponential scale ranging from 0.01 to 10. The additional $\alpha$ term for the Elastic-Net was tuned over 10 equa-distant values from 0 to 1.

@@

   4.	SMOTE (Synthetic Minority Over-Sampling Technique): This method incorporated both random under- and over-sampling. Under-sampling is performed without replacement from the majority class, while over-sampling is performed by creating new synthetic observations using the minority class. The SMOTE algorithm was devised by Chawla et al. [-@chawla2002smote] and stored in the DMwR package, available in R. The user is able to specify the parameters denoting the percent of over- and under-sampling from the two classes. This allows the user to control the proportion of disbalance in the data. For the purposes of this paper, we used SMOTE to create perfectly balanced data. However further research would be needed to see if varying the proportion of class disbalance could improve predictive performance.

###Classification

A classification algorithm in the form of a random forest was performed and validated using the test set. 10-fold cross validation was used to reduced bias due to random dataset generation. The default number of features to consider at each node of the random forest algorithm was set as the square root of the number of features in the model. Likewise, the number of ensemble trees to aggregate was set at 500. Once the model was implemented, it was assessed using the test set, and performance metrics were then recorded.

@@

###Model Evaluation

The efficacy of different ensemble systems was compared using various performance metrics including accuracy, AUC, F1-score, and MCC. These metrics are defined as follows:

$$
\begin{equation}
\begin{aligned}
Accuracy &= \dfrac{TP + TN}{TP + TN + FP + FN} \\
F1 - Score &= \dfrac{2TP}{2TP + FP + FN} \\
MCC &= \dfrac{TP \times TN - FP \times FN}{\sqrt{ (TP + FP)(TP+FN)(TN+FP)(TN+FN) }}
\end{aligned}
\end{equation}
$$
Here, TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary. Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. Accuracy measures the percentage of correct classifications by the model. However, in the case of heavily imbalanced class, as can be seen by the numerator, the metric is biased toward correcly identified majority classes. The F1-score and MCC are more balanced measures. We also considered the AUC for each model, by computing the average trapezoidal approximations for the curve created by the true positive rates (TPR) and false positive rates (FPR)
